
# About the course 

This course is about how to process open data with R. 

# Chapter 1 Setting up 

## Summary
- Installed RStudio, Github 
- Learned the basic R codes from DataCamp  "R Short and Sweet"
- Github
    + Fork a project from: https://github.com/TuomoNieminen/IODS-project. This will go to my own repository under my own profile
    + enabled a web page which combines my weekly exercises (index.html). This is my course diary web page
    + make changes to the local copy and push the changes into remote GitHub repository. "Commit to master" often and even on small changes
    + finally, sync

## R Markdown

When I click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. I can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

I can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Tables

First Header  | Second Header
------------- | -------------
Content Cell  | Content Cell
Content Cell  | Content Cell

Blockquotes

A friend once said:

> The only thing can stop a bad guy with a gun
> is a good guy with a gun



